---
title: "LAGOS Spatial Analysis"
author: "Matthew Ross"
date: "9/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
lagosne_get(dest_folder= LAGOSNE::lagos_path())

#once downloaded, need load into R
lagos <- lagosne_load() #leaving defaults can be best for newest data

names(lagos)
```

We are going to pull in lake centroids

```{r}
lake_centers <- lagos$locus
names(lake_centers)
```


### Convert to spatial data
# st_as_sf converts df to spatial
# (lats are x, long y) and need projection

```{r}
spatial_lakes <- st_as_sf(lake_centers, coords = 
                            c('nhd_long', 'nhd_lat'), crs = 4326) %>% # here we are telling r the current projection 
  st_transform(2163) # and here is the desired projection

subset_spatial <- spatial_lakes %>% 
  slice(1:100)

mapview(subset_spatial)
```

### Subset to only Minnesota

```{r}
states <- us_states() #from USAbounaries

minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

mapview(minnesota)
```

### Subset to only Minnesota

```{r}
#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```

# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}

the_ayes <- states %>%
  filter(state_name == c('Iowa', 'Illinois')) %>%
  st_transform(2163)

mapview(the_ayes)
```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

```{r}

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}


```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}

```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 


